---
title: "symbols layer for ggplot2"
author: |
  | Shuangbin Xu
  | School of Basic Medical Sciences, Southern Medical University
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    toc: true
    theme: cayman
    highlight: vignette
  pdf_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{ ggsymbol: basics using.}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, results="asis", message=FALSE, KnitrSetUp}
knitr::opts_chunk$set(tidy=FALSE,warning=FALSE,message=FALSE)
Biocpkg <- function (pkg){
    sprintf("[%s](http://bioconductor.org/packages/%s)", pkg, pkg)
}

CRANpkg <- function(pkg){
    cran <- "https://CRAN.R-project.org/package"
    fmt <- "[%s](%s=%s)"
    sprintf(fmt, pkg, cran, pkg)
}
```

```{r, echo=FALSE, results="hide", message=FALSE, Loadpackages}
library(ggplot2)
library(ggsymbol)
```

# 1. Introduction

The point shapes of [`ggplot2`](https://ggplot2.tidyverse.org/) based on `grid` provide several symbols. And the shapes are stored internally as integers. You can see the `pch` of `points`. Recently. I developed [`ggstar`](https://github.com/xiangpin/ggstar) to create some regular polygons (Non-deformable). It is built with the merits of [`ggplot2`](https://ggplot2.tidyverse.org/) that was based on the grammar of graphics. It can use `starshape` parameter to control the selection of shapes. However, It can not be mapped with `geom_point`. To fill the gap. The [`ggsymbol`](https://github.com/xiangpin/ggsymbol) was born. It provides points layer by integrating the shapes of `geom_point` of [`ggplot2`](https://ggplot2.tidyverse.org/) and some `starshapes` of [`ggstar`](https://github.com/xiangpin/ggstar)

# 2. Total shapes

The total shapes:

```{r, fig.align="center", fig.width=7, fig.height=5.5, showshapes}
library(ggplot2)
library(ggsymbol)
d <- data.frame(p=c(0:127),f=c(rep("g",26), rep("s", 7), rep("g", 95)))
d$f <- factor(d$f, levels=c("g", "s"))
p <- ggplot() + 
     geom_symbol(data=d, 
                 mapping=aes(x=p%%16, y=p%/%16, symbol=p, fill=f), 
                 size=4, stroke=0.5, show.legend=FALSE) +
     geom_text(data=d, 
               mapping=aes(x=p%%16, y=p%/%16+0.25, label=p), 
               size=3) +
     scale_symbol_identity() +
     scale_fill_manual(values=c("red", "blue")) +
     xlab(NULL) +
     ylab(NULL) +
     theme(axis.ticks=element_blank(),
           axis.text=element_blank())
p
```

The shapes of `0-25` and `33-127` are from `pointsGrob` of `grid`, the `26-32` are from `starGrob` of [`ggstar`](https://github.com/xiangpin/ggstar)

# 3. Demo

```{r, fig.align="center", fig.width=7, fig.height=5.2, Demo}
p <- ggplot(data=iris, aes(x=Sepal.Width,y=Sepal.Length)) +
     geom_symbol(aes(symbol=Species, fill=Species), 
                 color="grey60", size=2.5,
                 stroke=0.5) +
     scale_fill_manual(values=c("#E41A1C", "#377EB8", "#4DAF4A")) +
     scale_symbol_manual(values=c(28, 22, 21))

p
```

# 4. Need helps?

If you have questions/issues, please visit [github issue tracker](https://github.com/xiangpin/ggsymbol/issues).

# 5. Session information

Here is the output of sessionInfo() on the system on which this document was compiled:

```{r, echo=FALSE}
sessionInfo()
```

